### 2017-02-25 draft-md-ct.tex
several changes explained below
Main change:
Section about the continuous time analysis cleaned following [1].
Also there is a new section about the discrete time analysis also following [1].

Another change:
Switched to using $x'$ instead of $x^*$ for an arbitrary "competing" vector in the opening section.
Reason: want to reserve the "star" for duality.

% ==================================================
% ========== Draft: Mirror Descent in Continuous Time
% ==================================================
% Last modified: 2017 February 01

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amscd,amsthm}
\usepackage{mathdef-omar}
\usepackage{enumitem}

% ========== Pstricks
%\usepackage{pst-all}

% ========== Page formats
\setlength{\oddsidemargin}{1.5cm}
\setlength{\textwidth}{14cm}
\setlength{\topmargin}{-.5cm}
\setlength{\textheight}{22cm}

% ========== Theorem-like stuff
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{coro}{Corollary}
\newtheorem{propo}{Proposition}
\newtheorem*{claim}{Claim}
%
\theoremstyle{definition}
\newtheorem{defi}{Definition}
\newtheorem{example}{Example}
%
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{remarks}{Remarks}
\newtheorem*{note}{Note}
\newtheorem*{notes}{Notes}
\newtheorem*{notation}{Notation}
\newtheorem*{notations}{Notations}
%
\newtheorem{fact}[theorem]{Fact}
\newtheorem{exercise}[theorem]{Exercise}
%
\renewcommand{\proof}{\noindent {\sc Proof.}\quad}
\renewcommand{\qedsymbol}{$\blacksquare$}
% ==========

\begin{document}  % ========== BEGIN DOCUMENT

\section*{Some calculations about the MD algorithm}

\noindent
The setting and notations...
\begin{itemize}[leftmargin=*, itemsep=2pt]
\item The potential $R : \R^d \to \R$, a Legendre function.
\item Its convex conjugate: $R^{*}(u) = \sup_{x} \{ \ab{u,x} - R(x)\}$.
\item The ``mirror maps'' $\nabla R : \R^d \to \R^d$ and $\nabla R^{*} : \R^d \to \R^d$.
\item The loss functions $(\ell_t)$, each $\ell_t : \R^d \to \R$.
\item Time index $t \in \{ 0,1,2,\ldots \}$ for discrete time, $t \in [0,\infty)$ for continuous time.
\item $x$, $x_t$ (discrete time), $x(t)$ (continuous time) elements of the primal space.
\item $u$, $u_t$ (discrete time), $u(t)$ (continuous time) elements of the dual space.
\end{itemize}
The MD algorithm (discrete time) uses the initialization and updates
\begin{align}
x_1 &= \arg\min R(x)\,, \label{eq.dtime.md1} \tag{1md} \\
x_{t+1} &= \arg\min_{x \in \R^d} \Bigl\{ \eta \ab{\nabla\ell_t(x_{t}), x} + D_{R}(x, x_{t})\Bigr\}\,.
 \label{eq.dtime.md2} 
\tag{2md}
\end{align}
Here $D_{R}(x,y)$ stands for the Bregman divergence, whose definition is recalled now:
\[
D_{R}(a_{1}, a_{2}) = R(a_{1}) - R(a_{2}) - \ab{\nabla R(a_2), a_{1} - a_{2}}\,.
\]
The order of the arguments matters.
Let's say $D_{R}(a_{1},a_{2})$ is ``supported at $a_{2}$''
(its definition uses a tangent supported at $a_{2}$).
To minimize clutter let's denote by $\partial_{i}$ the partial derivative (operator)
with respect to the $i$-th argument. So for instance:
\begin{align*}
\partial_{1} D_{R}(a_{1}, a_{2}) &= \nabla R(a_{1}) - \nabla R(a_{2})\,, \\[1mm]
\partial_{2} D_{R}(a_{1}, a_{2}) &= - \nabla^2 R(a_{2})(a_{1} - a_{2})\,.
\end{align*}
The equations \eqref{eq.dtime.md1}--\eqref{eq.dtime.md2} are equivalent to:
\begin{align*}
u_{t+1} &= u_{t} - \eta \nabla \ell_{t}(x_{t}) \label{eq.dtime.1} \tag{1dt} \\
x_{t} &= \nabla R^{*}(u_{t})\,. \label{eq.dtime.2} \tag{2dt}
\end{align*}
Analogous equations in continuous time:
\begin{align*}
\dot{u}(t) &= - \eta \nabla \ell_{t}(x(t)) \label{eq.ctime.1} \tag{1ct} \\
x(t) &= \nabla R^{*}(u(t))\,. \label{eq.ctime.2} \tag{2ct}
\end{align*}
%For a fixed $x^{*} \in \R^d$, $u^{*} = \nabla R(x^{*})$,
For a fixed $x' \in \R^d$, writing $u' = \nabla R(x')$ for its image in the dual space,
\setcounter{equation}{3}
\begin{equation}
%\ell_{t}(x^{*}) \geq \ell_{t}(x(t)) + \ab{\nabla\ell_t(x(t)), x^{*} - x(t)} \label{eq.linearizing.loss}
\ell_{t}(x') \geq \ell_{t}(x(t)) + \ab{\nabla\ell_t(x(t)), x' - x(t)}\,. \label{eq.linearizing.loss}
\end{equation}
From \eqref{eq.ctime.1} we get $\nabla \ell_{t}(x(t)) = -\dot{u}(t)/\eta$,
and plugging this in \eqref{eq.linearizing.loss} we get:
\[
%\ell_{t}(x^{*}) \geq \ell_{t}(x(t)) - \frac{1}{\eta} \ab{\dot{u}(t), x^{*} - x(t)}\,.
\ell_{t}(x') \geq \ell_{t}(x(t)) - \frac{1}{\eta} \ab{\dot{u}(t), x' - x(t)}\,.
\]
Rearranging:
\begin{equation}
%\ell_{t}(x(t)) - \ell_{t}(x^{*}) \leq  \frac{1}{\eta} \ab{\dot{u}(t), x^{*} - x(t)}\,, \label{eq.loss.prep}
\ell_{t}(x(t)) - \ell_{t}(x') \leq  \frac{1}{\eta} \ab{\dot{u}(t), x' - x(t)}\,, \label{eq.loss.prep}
\end{equation}
and the claim is that the RHS is the derivative of a Bregman divergence.

\bigskip

Recall the identity $D_{R^{*}}(u', u) = D_{R}(x,x')$, which holds for any `dual pairs'
$u = \nabla R (x)$, $u' = \nabla R (x')$.

\bigskip

On one hand, by the chain rule and the symmetry of the Hessian:
\begin{align*}
%\frac{d}{dt} D_{R}(x(t), x^{*})
%\frac{d}{dt} D_{R^{*}}(u^{*}, u(t))
%&= \ab{ \partial_{2} D_{R^{*}}(u^{*}, u(t)), \dot{u}(t)} \\[1mm]
%&= - \ab{ \nabla^2 R^{*}(u(t))(u^{*} -u(t)), \dot{u}(t) } \\[1mm]
%&= - \ab{ u^{*} -u(t), \nabla^2 R^{*}(u(t)) \dot{u}(t) } \\[1mm]
%&= - \ab{u^{*} - u(t), \dot{x}(t)}\,.
%
\frac{d}{dt} D_{R}(x(t), x')
=\frac{d}{dt} D_{R^{*}}(u', u(t))
&= \ab{ \partial_{2} D_{R^{*}}(u', u(t)), \dot{u}(t)} \\[1mm]
&= - \ab{ \nabla^2 R^{*}(u(t))(u' -u(t)), \dot{u}(t) } \\[1mm]
&= - \ab{ u' -u(t), \nabla^2 R^{*}(u(t)) \dot{u}(t) } \\[1mm]
&= - \ab{u' - u(t), \dot{x}(t)}\,.
\addtocounter{equation}{1}
\label{eq.deriv.diver.dual}  \tag{\theequation}
\end{align*}
On the other hand,
\begin{align*}
%\frac{d}{dt} \ab{u^{*} - u(t), x^{*} - x(t)}
%&= - \ab{ \dot{u}(t), x^{*} - x(t)} - \ab{ u^{*} -u(t), \dot{x}(t) }\,.
\frac{d}{dt} \ab{u' - u(t), x' - x(t)}
&= - \ab{ \dot{u}(t), x' - x(t)} - \ab{ u' -u(t), \dot{x}(t) }\,.
\end{align*}
Combining the two:
\begin{align*}
%\ab{ \dot{u}(t), x^{*} - x(t)}
%&= - \ab{ u^{*} -u(t), \dot{x}(t) } - \frac{d}{dt} \ab{u^{*} - u(t), x^{*} - x(t)} \\[1mm]
%%&= \frac{d}{dt} D_{R}(x(t), x^{*}) - \frac{d}{dt} \ab{u^{*} - u(t), x^{*} - x(t)}\,.
%&= \frac{d}{dt} \Bigl( D_{R^{*}}(u^{*}, u(t)) - \ab{u^{*} - u(t), x^{*} - x(t)} \Bigr)\,.
%
\ab{ \dot{u}(t), x' - x(t)}
&= - \ab{ u' -u(t), \dot{x}(t) } - \frac{d}{dt} \ab{u' - u(t), x' - x(t)} \\[1mm]
%&= \frac{d}{dt} D_{R}(x(t), x') - \frac{d}{dt} \ab{u' - u(t), x' - x(t)}\,.
&= \frac{d}{dt} \Bigl( D_{R}(x(t), x') - \ab{u' - u(t), x' - x(t)} \Bigr)\,.
\end{align*}
Note that the term
%$\ab{u^{*} - u(t), x^{*} - x(t)} = \ab{\nabla R(x^{*}) - \nabla R(x(t)), x^{*} - x(t)}$
%is the sum of $D_{R}(x^{*}, x(t))$ and $D_{R}(x(t), x^{*})$.
$\ab{u' - u(t), x' - x(t)} = \ab{\nabla R(x') - \nabla R(x(t)), x' - x(t)}$
is the sum of $D_{R}(x', x(t))$ and $D_{R}(x(t), x')$.
This is a general property of Bregman divergences that can be obtained directly from the definition:
$$
D_{R}(a_{1}, a_{2}) + D_{R}(a_{2}, a_{1})
= \ab{\nabla R(a_{1}) - \nabla R(a_{2}), a_{1} - a_{2}}\,.
$$
Then \eqref{eq.loss.prep} can be written as
\begin{align*}
%\ell_{t}(x(t)) - \ell_{t}(x^{*})
%\leq  \frac{1}{\eta} 
%   \frac{d}{dt} \Bigl( D_{R}(x(t), x^{*}) - D_{R}(x^{*}, x(t)) - D_{R}(x(t), x^{*})\Bigr)\,.
%= -\frac{1}{\eta} \frac{d}{dt} D_{R}(x^{*}, x(t))\,.
%
\ell_{t}(x(t)) - \ell_{t}(x')
&\leq  \frac{1}{\eta} \frac{d}{dt} \Bigl( D_{R}(x(t), x') - D_{R}(x', x(t)) - D_{R}(x(t), x')\Bigr) \\[1mm]
&= -\frac{1}{\eta} \frac{d}{dt} D_{R}(x', x(t))\,.
\end{align*}
After integrating:
\begin{align*}
%&\int_{0}^{T} [\ell_{t}(x(t)) - \ell_{t}(x^{*})] dt \\
%&\hspace{1.5cm}
%\leq \frac{1}{\eta} \bigl[ D_{R^{*}}(u^{*}, u(T)) - D_{R}(x^{*}, x(T)) - D_{R}(x(T), x^{*}) \\
%&\hspace{3.0cm} 
%- D_{R^{*}}(u^{*}, u(0)) + D_{R}(x^{*}, x(0)) + D_{R}(x(0), x^{*}) \bigr]\,.
%
\int_{0}^{T} [\ell_{t}(x(t)) - \ell_{t}(x')] dt 
\leq \frac{ D_{R}(x', x(0)) - D_{R}(x', x(T)) }{\eta}\,.
\addtocounter{equation}{1}
\label{eq.bound.first}  \tag{\theequation}
\end{align*}

Note: Follow the regularized leader (FTRL) is equivalent to mirror descent when $R$ is a Legendre function
(unconstrained ``optimization'').
Hence, the above result is applicable to both MD and FTRL in this case.

However, when $R$ is not Legendre (constrained case), they will generally be different.

\medskip

\section*{Equivalence between FTRL and MD}

Some notes to clarify the equivalence between 
Follow the Regularized Leader (FTRL) and Mirror Descent (MD) algorithms.

\medskip

The FTRL algorithm (discrete time) uses the initialization and updates
\begin{align}
x_1 &= \arg\min R(x) \label{eq.ftrl1} \tag{1ftrl} \\
x_{t+1} &= \arg\min_{x \in \R^d} \Bigl\{ \eta \Bigab{ \sum_{s=1}^{t} \nabla\ell_s(x_{s}), x} + R(x)\Bigr\}\,.
 \label{eq.ftrl2} 
\tag{2ftrl}
\end{align}
Here $R(x)$ is the regularizer used by FTRL.
To find an explicit form for the optimizer, let
$$
G(x) = \eta \sum_{s=1}^{t} \ab{ \nabla\ell_s(x_{s}), x} + R(x)\,.
$$
Then $\nabla G(x) = \eta \sum_{s=1}^{t} \nabla\ell_s(x_{s}) + \nabla R(x)$
and it is clear that $\nabla G(x) = 0$ 
when $x = \nabla R^{*} \bigl( -\eta \sum_{s=1}^{t} \nabla\ell_s(x_{s}) \bigr)$.
The second derivative $\nabla^2 G(x) = \nabla^2 R(x)$ is positive,
so this point $x$ gives the minimum.
So the update rule \eqref{eq.ftrl2} becomes
$$
x_{t+1} = \nabla R^{*} \bigl( -\eta \sum_{s=1}^{t} \nabla\ell_s(x_{s}) \bigr)\,.
$$
Defining $u_{t+1} = -\eta \sum_{s=1}^{t} \nabla\ell_s(x_{s})$ with $u_{0} = 0$,
we get $u_{t+1} = u_{t} - \eta \nabla\ell_t(x_{t})$,
and the update can be rewritten as the following:
\begin{align*}
u_{t+1} &= u_{t} - \eta \nabla \ell_{t}(x_{t})\,, \\
x_{t+1} &= \nabla R^{*}(u_{t+1})\,. 
\end{align*}
These are clearly the equations \eqref{eq.dtime.1}--\eqref{eq.dtime.2} of MD.

\medskip

Conversely, starting from the equations \eqref{eq.dtime.1}--\eqref{eq.dtime.2} of MD,
we can revert the process and obtain the equations \eqref{eq.ftrl1}--\eqref{eq.ftrl2} of FTRL.

\medskip

\section*{Continuous time analysis}

Based on the approach of \cite[Section 4]{Kwon-M}.
For the time being using a notation closer to theirs but not conflicting with ours.
Keeping a constant learning rate $\eta > 0$, their process looks like this:
\begin{align*}
x(t) = \nabla R^{*} \Bigl( \eta \int_{0}^{t} v(s) ds \Bigr).
\end{align*}
In this setting, $v(s) = -\nabla \ell_{s}(x(s))$ in the case of differentiable loss functions,
or more generally $v(s) \in \partial (-\ell_{s}(x(s)))$ is a `negative sub-gradient'
when only convexity but no differentiability is assumed on the loss functions.

\medskip

Setting $y(t) = \eta \int_{0}^{t} v(s) ds$, the above process can be presented as
\begin{align*}
\dot{y}(t) &= \eta v(t) \label{eq.cta.1} \tag{1ct'} \\
x(t) &= \nabla R^{*}(y(t))\,. \label{eq.cta.2} \tag{2ct'}
\end{align*}
Then for a fixed $x$,
\begin{equation}
\int_{0}^{t} \ab{v(s), x} ds
= \frac{\ab{y(t), x}}{\eta}
\leq \frac{R^{*}(y(t)) + R(x)}{\eta}\,.
\label{eq.cta.using.fenchel.ineq}
\end{equation}
But noting from \eqref{eq.cta.2} that $\nabla R (x(t)) = y(t)$, 
which means that $x(t)$ and $y(t)$ satisfy equality in the Fenchel inequality,
then
\begin{align}
\frac{R^{*}(y(t))}{\eta}
= \frac{\ab{y(t), x(t)} - R(x(t))}{\eta} 
= \int_{0}^{t} \ab{v(s), x(t)} ds - \frac{R(x(t))}{\eta}\,.
\label{eq.cta.something}
\end{align}
Consider the function $\varphi(t,x) = \int_{0}^{t} \ab{v(s), x} ds - R(x)/\eta$.
Its partial derivatives are $\partial_{t} \varphi(t,x) = \ab{v(t),x}$ and
$\partial_{x} \varphi(t,x) = y(t) - \nabla R(x)/\eta$.
Replacing $x$ with $x(t)$ we have 
$$
\varphi(t,x(t)) = \int_{0}^{t} \ab{v(s), x(t)} ds - \frac{R(x(t))}{\eta}\,.
$$
The time derivative of this is
\begin{align*}
\frac{d}{dt}\varphi(t,x(t))
= \partial_{t}\varphi(t,x(t)) + \ab{\partial_{x} \varphi(t,x(t)), \dot{x}(t)}
= \ab{v(t), x(t)}\,.
\end{align*}
Notice that $\partial_{x} \varphi(t,x(t)) = y(t) - \nabla R(x(t))/\eta = 0$ because of \eqref{eq.cta.2}.
Using this in \eqref{eq.cta.something}:
\begin{align*}
\frac{d}{dt} \frac{R^{*}(y(t))}{\eta}
= \frac{d}{dt}\varphi(t,x(t))
= \ab{v(t), x(t)}\,,
\end{align*}
and integrating:
\begin{align*}
\frac{R^{*}(y(t)) - R^{*}(y(0))}{\eta}
= \int_{0}^{t} \ab{v(s), x(s)} ds\,.
\end{align*}
Then, continuing from \eqref{eq.cta.using.fenchel.ineq},
and rearranging, we get
\begin{align*}
\int_{0}^{t} \ab{v(x), x} ds - \int_{0}^{t} \ab{v(s), x(s)} ds 
\leq \frac{R(x) + R^{*}(y(0))}{\eta}\,.
\addtocounter{equation}{1}
\label{eq.cta.bound.shaping}  \tag{\theequation}
\end{align*}
Note that $R^{*}(y(0)) = R^{*}(0) = - R_{\text{min}}$. 
Finally, using the linearization trick (see \eqref{eq.linearizing.loss} above),
and recall that $v(s) \in \partial (-\ell_{s}(x(s)))$ is a negative sub-gradient,
we get
\begin{align*}
\int_{0}^{t} [\ell_{s}(x(s)) - \ell_{s}(x)] ds
\leq \int_{0}^{t} \ab{v(s), x - x(s)} ds 
\leq \frac{R(x) - R_{\text{min}}}{\eta}\,.
\end{align*}
Of course we could upper bound $R(x)$ by $R_{\text{max}}$ which
would give us the exactly the upper bound of \cite{Kwon-M}:
\begin{align*}
\int_{0}^{t} [\ell_{s}(x(s)) - \ell_{s}(x)] ds
\leq \int_{0}^{t} \ab{v(s), x - x(s)} ds 
\leq \frac{R_{\text{max}} - R_{\text{min}}}{\eta}\,.
\addtocounter{equation}{1}
\label{eq.cta.bound}  \tag{\theequation}
\end{align*}

\medskip

\section*{Discrete time analysis}

Continuing as in \cite[Section 5]{Kwon-M}.
Consider the discrete-time updates
\begin{align*}
y_{t+1} &= y_{t} - \eta v_{t}\,, \label{eq.dta.1} \tag{1dt'} \\
x_{t+1} &= \nabla R^{*}(y_{t+1})\,. \label{eq.dta.2} \tag{2dt'}
\end{align*}
These are the discrete-time counterparts of \eqref{eq.cta.1}--\eqref{eq.cta.2} form
the previous section. As in \cite{Kwon-M}, let's start by embedding the discrete-time process
$(v_{t})_{t \geq 0}$ into a continuous-time stream $v(t)$ by setting
\begin{equation}
v(t) = v_{\lceil t \rceil},\hspace{7mm} t \in [0,\infty)\,.
\label{eq.dt.to.ct.embed}
\end{equation}
We will derive a regret bound for the strategy \eqref{eq.dta.1}--\eqref{eq.dta.2}.
The bound will consist of a term coming form the continuous time analysis of the previous section 
applied to the continuous stream \eqref{eq.dt.to.ct.embed}, 
plus a term accounting for a comparison between the discrete and continuous time processes.

\medskip

Assume we have a $K$-strongly convex regularizer $R$.
Denoting as before $x(t) = \nabla R^{*} (y(t))$ and $y(t) = \eta \int_{0}^{t} v(s) ds$,
we have
$$
x_{t} = \nabla R^{*} \Bigl( \eta \sum_{k=1}^{t-1} v_{k} \Bigr) = x(t-1).
$$
Note that $x_t$ on the left is the value of the discrete-time process at stage $t$,
while $x(t-1)$ in the right is the continuous-time process at time $t-1$.

\medskip

For any discrete time $k \geq 1$ and for any continuous time $t \in (k-1,k)$,
the difference between the (linearized) losses is upper bounded as follows:
\begin{align*}
\abs{ \ab{v(t),x(t)} - \ab{v_{k},x_{k}} }
&= \abs{ \ab{v_{k}, x(t) - x(k-1)} } \\[1.5mm]
&\leq \norm{v_{k}}_{*} \norm{ \nabla R^{*}(y(t)) - \nabla R^{*}(y(k-1)) } \\
&\leq \frac{1}{K} \norm{v_{k}}_{*} \norm{y(t) - y(k-1)}_{*}\,.
\end{align*}
The last step used that $\nabla R^{*}$ is $(1/K)$-Lipschitz,
which is a consequence of (in fact equivalent to) the $K$-strong convexity of $R$.
But the last term can be bounded as follows:
$$
\norm{y(t) - y(k-1)}_{*}
= \Bignorm{ \eta \int_{k-1}^{t} v(s) ds }
\leq \eta (t-k+1) \norm{v(k)}_{*}\,.
$$
Therefore, we get the promised comparison between the discrete and the continuous time
(linearized) losses:
\begin{align*}
\Bigabs{ \int_{0}^{t} \ab{v(s),x(s)} ds - \sum_{k=1}^{t} \ab{v_{k},x_{k}} }
&\leq \sum_{k=1}^{t} \int_{k-1}^{k} \bigabs{ \ab{v(s),x(s)} - \ab{v_{k},x_{k}} } ds \\
&\leq \frac{\eta}{K} \sum_{k=1}^{t} \norm{v_{k}}_{*} \norm{v(k)}_{*} \int_{k-1}^{k} (s-k+1) ds \\
&= \frac{\eta}{2K} \sum_{k=1}^{t} \norm{v_{k}}_{*}^2\,.
\end{align*}

Now, for a fixed arbitrary $x$, we get
\begin{align*}
\sum_{k=1}^{t} \ab{v_{k},x}
= \int_{0}^{t} \ab{v(s), x} ds 
&\leq \int_{0}^{t} \ab{v(s), x(s)} ds + \frac{R_{\text{max}} - R_{\text{min}}}{\eta} \\
&\leq \sum_{k=1}^{t} \ab{v_{k},x_{k}} + \frac{\eta}{2K} \sum_{k=1}^{t} \norm{v_{k}}_{*}^2 
     + \frac{R_{\text{max}} - R_{\text{min}}}{\eta}\,.
\end{align*}
The first inequality is from \eqref{eq.cta.bound} and the second inequality
from the above continuous/discrete time comparison for the linearized losses.
Finally, using linearization of the original losses (in discrete time) we get
\begin{align*}
\sum_{k=1}^{t} \ell_{k}(x_{k}) - \sum_{k=1}^{t} \ell_{k}(x)
&\leq \sum_{k=1}^{t} \ab{v_{k},x - x_{k}} \\
&\leq \frac{R_{\text{max}} - R_{\text{min}}}{\eta} + \frac{\eta}{2K} \sum_{k=1}^{t} \norm{v_{k}}_{*}^2\,.
\end{align*}

In particular, is $\norm{v_{k}}_{*} \leq M$ for all $k$ (and some $M > 0$),
then optimizing over $\eta$ we get a regret bound that looks like
$$
\sqrt{2M^2(R_{\text{max}} - R_{\text{min}}) \ t/K}\,.
$$

\medskip

\section*{Another continuous time analysis (?)}

Kind of based on the approach of \cite{Kwon-M} too, but some modifications.
This is the stuff I had written before, keeping it here for the time being just because.

\bigskip

\noindent
From \eqref{eq.ctime.1} we have
$$
u(t) - u(0) = \int_{0}^{t} \dot{u}(s) ds = -\eta \int_{0}^{t} \nabla\ell_{s}(x(s)) ds.
$$
Then for a fixed $x$,
\begin{equation}
\int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x} ds
= \frac{\ab{u(t) - u(0), x}}{\eta}
\leq \frac{R^{*}(u(t)) + R(x) - \ab{u(0),x}}{\eta}\,.
\label{eq.using.fenchel.ineq}
\end{equation}
But noting from \eqref{eq.ctime.2} that $x(t) = \nabla R^{*}(u(t))$, 
which means that $x(t)$ is the optimizer in the definition of $R^{*}$, 
then
\begin{align}
\frac{R^{*}(u(t))}{\eta}
&= \frac{\ab{u(t), x(t)} - R(x(t))}{\eta} \notag \\[1mm]
&= \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(t)} ds + \frac{\ab{u(0),x(t)}-R(x(t))}{\eta}\,.
\label{eq.something}
\end{align}
Consider the function $\varphi(t,x) = \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x} ds$.
In particular, replacing $x$ with $x(t)$ we have
$$
\varphi(t,x(t)) = \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(t)} ds.
$$
The time derivative is
\begin{align*}
\frac{d}{dt}\varphi(t,x(t))
%&= \dot{\varphi}(t,x(t)) + \frac{\partial}{\partial x} \varphi(t,x(t)) \cdot \dot{x}(t) \\
&= \ab{-\nabla\ell_{t}(x(t)), x(t)} + \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), \dot{x}(t)} ds \\
&= \ab{-\nabla\ell_{t}(x(t)), x(t)} + \frac{1}{\eta} \ab{u(t) - u(0), \dot{x}(t)} \\
&= \ab{-\nabla\ell_{t}(x(t)), x(t)} + \frac{1}{\eta} \frac{d}{dt} D_{R}(x(t), x(0)) \,,
\end{align*}
where in the last step we used \eqref{eq.deriv.diver.dual} with $x(0)$ instead of $x'$.
Then integrating
(note that $\varphi(0,x) = 0$ for any $x$, and $D_{R}(x(0), x(0)) = 0$) 
we get
\begin{align*}
\varphi(t,x(t)) 
&= \int_{0}^{t} \frac{d}{ds} \varphi(s,x(s)) ds \\
&= \int_{0}^{t} \ab{-\nabla\ell_{s}(x(s)), x(s)} ds 
     + \frac{1}{\eta} D_{R}(x(t), x(0)) \,.
\end{align*}
Then, continuing from \eqref{eq.using.fenchel.ineq} and \eqref{eq.something},
and rearranging, we get
\begin{align*}
&\int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x(s)} ds - \int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x} ds \\
&\hspace{1.5cm}
\leq \frac{1}{\eta} \Bigl[ R(x) - R(x(t)) - \ab{u(0), x - x(t)} + D_{R}(x(t), x(0)) \Bigr]\,.
\addtocounter{equation}{1}
\label{eq.bound.shaping}  \tag{\theequation}
\end{align*}
Note that $u(0) = \nabla R(x(0))$. We have
\begin{align*}
&R(x) - R(x(t)) - \ab{u(0), x - x(t)} \\[1mm]
&\hspace{1.5cm}
= R(x) - R(x(t)) - \ab{\nabla R(x(0)), x - x(t)} \\[1mm]
&\hspace{1.5cm}
= R(x) - R(x(0)) - \ab{\nabla R(x(0)), x - x(0)} \\
&\hspace{2.5cm}
   + R(x(0)) - R(x(t)) - \ab{\nabla R(x(0)), x(0) - x(t)} \\[1mm]
&\hspace{1.5cm}
= D_{R}(x,x(0)) + D_{R}(x(0),x(t))\,.
\end{align*}
Using this in \eqref{eq.bound.shaping} we get
\begin{align*}
&\int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x(s)} ds - \int_{0}^{t} \ab{\nabla\ell_{s}(x(s)), x} ds \\
&\hspace{1.5cm}
\leq \frac{1}{\eta} \Bigl[ D_{R}(x,x(0)) + D_{R}(x(0),x(t)) + D_{R}(x(t), x(0)) \Bigr]\,.
\end{align*}
Finally, using the linearization trick (see \eqref{eq.linearizing.loss} above), we get
\begin{align*}
&\int_{0}^{T} [\ell_{t}(x(t)) - \ell_{t}(x)] dt
\leq \int_{0}^{T} \ab{\nabla\ell_{t}(x(t)), x(t) - x} dt \\
&\hspace{2.0cm}
\leq \frac{1}{\eta} \Bigl[ D_{R}(x,x(0)) + D_{R}(x(0),x(T)) + D_{R}(x(T), x(0)) \Bigr]\,.
\addtocounter{equation}{1}
\label{eq.bound.new}  \tag{\theequation}
\end{align*}

\medskip

\noindent
Note: Comparing, the first bound \eqref{eq.bound.first} clearly beats this
`new' bound \eqref{eq.bound.new}.

\medskip

\begin{thebibliography}{biblio}

\bibitem{Kwon-M}
  J.~Kwon and P.~Mertikopoulos,
  A continuous-time approach to online optimization.
  Preprint. arXiv:1401.6956.

\bibitem{Warmuth-Jagota}
  M.K.~Warmuth and A.K.~Jagota,
  Continuous and discrete-time nonlinear gradient descent: relative loss bounds and convergence.
  In: 
  \textit{Proc. Fifth International Symposium on Artificial Intelligence and Mathematics}, 1998.

\end{thebibliography}

\vspace{.5cm}

% -------------------------------------------------------
\noindent
Last updated: \today

\end{document}  % ========== END DOCUMENT

%% Picture 
%\psset{unit=1.4pt} \vspace{3mm} \hspace{8cm}
%\pspicture(0,0)(0,0)
%\pspolygon[fillcolor=gray,fillstyle=crosshatch*](0,0)(70,0) (70,50)(0,50)
%\rput(35,25){\psframebox*[framearc=.3]{\parbox[c]{2.5cm}{\footnotesize\centering STEEL PLATE ON ROAD}}}
%\endpspicture